\section*{Frontier Worklog}

\subsection*{2026-06-18 Paper Plot Summary Tables}

\begin{itemize}
\item Treated \texttt{\detokenize{frontier_decoder2.tex}} as the current paper source for the figure inventory, with sha256 \texttt{\detokenize{d1abc814aab7ec6e8bacfab0af31b95d7f84b4a01170b648c2ceefacd5ae153e}}.
\item Added compact plot-ready CSV tables and JSON sidecars for every current figure or panel, including the generated schematic, BB72 algorithm recap state table, surface/color code-capacity panels, surface memory-Z DEM comparison, BB72/Gross detector-side DEM panels, Gross transition-evaluation tail/percentile data, and the Gross \(p=0.002\) failure-decomposition table.
\item Left all manifest rows as \texttt{\detokenize{script-missing}}, not \texttt{\detokenize{reproducible}}, because the committed repo still lacks figure-specific renderers for the published PNGs.
\item Normalized local and scratch source paths in imported CSVs to stable source labels and avoided copying raw per-shot corpora.
\item Updated paper-plot docs and tests so \texttt{\detokenize{script-missing}} rows must have committed CSV data, same-stem sidecars, and matching checksums while the CLI continues to skip unreproducible rows honestly.
\end{itemize}

\subsection*{2026-06-18 Paper Plot Reproduction Scaffold}

\begin{itemize}
\item Audited the checkout for paper figure lists and plot-ready summary tables; none were present, so no paper plot data or numeric figure values were added.
\item Added an honest paper plot scaffold under \texttt{\detokenize{paper/plots/}}: schema-only \texttt{\detokenize{manifest.csv}}, missing-data documentation, ignored generated outputs, and \texttt{\detokenize{reproduce_plots.py}} commands that list or skip missing rows without fabricating data.
\item Extended \texttt{\detokenize{tools/asset_manifest.py}} with optional \texttt{\detokenize{--root}} and \texttt{\detokenize{--title}} arguments while preserving the default Gross/BB144 asset manifest output, then added \texttt{\detokenize{paper/plots/data/MANIFEST.md}} for the current retained paper-data directory.
\item Added \texttt{\detokenize{tests/test_paper_plots.py}} so the manifest schema, missing-data honesty, plot reproduction entry point, and data-manifest checksums are enforced.
\item Added \texttt{\detokenize{codemeta.json}}, \texttt{\detokenize{AUTHORS.md}}, \texttt{\detokenize{SECURITY.md}}, and \texttt{\detokenize{constraints/py312-ubuntu-ci.TODO.md}} using only declared repository facts, and updated README, AGENTS, CONTRIBUTING, reproducibility, release, metadata, file-scope, and typecheck docs accordingly.
\end{itemize}

\subsection*{2026-06-18 Metadata Finalization}

\begin{itemize}
\item Replaced provisional metadata with declared current status: software authors Anthony Leverrier and Rüdiger Urbanke, no declared ORCIDs, no assigned DOI, Plan France 2030 project ANR-22-PETQ-0006 funding, COSMIQ/Inria sabbatical hospitality acknowledgement, and no separate asset-license statement beyond the repository \texttt{\detokenize{NOTICE}}.
\item Added \texttt{\detokenize{constraints/py314-macos-validated.txt}} from the MacBook Python 3.14.2 validation environment and removed the provisional constraints note.
\item Removed provisional markers from public docs and the depth-8 backend error message, then added a regression test that checks public docs stay free of those markers.
\end{itemize}

\subsection*{2026-06-18 Academic Release Metadata Pass}

\begin{itemize}
\item Added public research-software metadata and release hygiene files: \texttt{\detokenize{CITATION.cff}}, \texttt{\detokenize{ACKNOWLEDGEMENTS.md}}, \texttt{\detokenize{CONTRIBUTING.md}}, \texttt{\detokenize{CHANGELOG.md}}, \texttt{\detokenize{docs/ACADEMIC_METADATA.md}}, \texttt{\detokenize{docs/ASSET_PROVENANCE.md}}, \texttt{\detokenize{docs/REPRODUCIBILITY.md}}, \texttt{\detokenize{docs/RELEASE.md}}, and \texttt{\detokenize{constraints/}}.
\item Added \texttt{\detokenize{tools/asset_manifest.py}} and generated \texttt{\detokenize{docs/ASSET_MANIFEST.md}} with SHA256 checksums for all bundled Gross/BB144 assets.
\item Updated README navigation, preferred public API guidance, architecture docs, file-scope audit, licensing notes, Makefile, CI, and tests for the publication-readiness checklist.
\item Modernized \texttt{\detokenize{pyproject.toml}} to the current PyPA license expression form: \texttt{\detokenize{license = "Apache-2.0"}} plus \texttt{\detokenize{license-files = ["LICENSE", "NOTICE"]}}.
\item Added exact dependency constraints for the validated MacBook Python 3.14.2 environment in \texttt{\detokenize{constraints/py314-macos-validated.txt}}.
\item Validation completed in a temporary Python 3.14 virtualenv: editable install, native build, \texttt{\detokenize{pytest -q}} with \texttt{18 passed}, frontier smoke, \texttt{\detokenize{rotated_surface_d3}} DEM info, \texttt{\detokenize{bravyi_depth7}} DEM info, both Python examples, the rotated-surface replay shell example, regenerated-manifest diff check, private-path grep, and \texttt{\detokenize{git diff --check}}.
\end{itemize}

\subsection*{2026-06-18 Apache-2.0 Licensing}

\begin{itemize}
\item Added the root \texttt{\detokenize{LICENSE}} file with the standard Apache License 2.0 text and a root \texttt{\detokenize{NOTICE}} file for repository attribution plus bundled-asset provenance notes.
\item Updated \texttt{\detokenize{pyproject.toml}} with \texttt{\detokenize{license = \{ file = "LICENSE" \}}} and the Apache Software License classifier.
\item Replaced the former license-status note in \texttt{\detokenize{docs/LICENSING.md}} with Apache-2.0 scope guidance, third-party licensing notes, and the rationale for not using CC BY-NC-SA as a software-code license.
\item Updated README navigation and the README license section, plus \texttt{\detokenize{docs/ARCHITECTURE.md}}, \texttt{\detokenize{AGENTS.md}}, and \texttt{\detokenize{docs/FILE_SCOPE.md}}, so humans and agents preserve Apache-2.0 and third-party notices in future changes.
\item Added a lightweight regression test that checks the license text, docs, README, package metadata, and file-scope audit.
\end{itemize}

\subsection*{2026-06-18 Navigation Polish and CLI Testability}

\begin{itemize}
\item Removed developer-machine absolute paths from validation notes and kept interpreter references generic as \texttt{\detokenize{$PYTHON_BIN}}.
\item Added \texttt{\detokenize{frontier.progressive}} as a public wrapper for small model-construction primitives, then updated \texttt{\detokenize{examples/minimal_decode.py}} to use only public \texttt{\detokenize{frontier.*}} imports.
\item Removed \texttt{\detokenize{sys.path}} manipulation from Python examples; they now model installed editable-package usage.
\item Normalized all CLI modules around \texttt{\detokenize{main(argv=None)}} and \texttt{\detokenize{_parse_args(argv)}}, including \texttt{\detokenize{tools.dem_loader}}, \texttt{\detokenize{tools.frontier_decoder}}, \texttt{\detokenize{tools.frontier_sample_replay}}, and \texttt{\detokenize{tools.frontier_bb144_benchmark}}.
\item Corrected \texttt{\detokenize{docs/ENVIRONMENT.md}} to match \texttt{\detokenize{resolve_cache_root}}, fixed the rotated-surface replay command label in \texttt{\detokenize{docs/COMMANDS.md}}, and documented platform/compiler expectations in \texttt{\detokenize{docs/ARCHITECTURE.md}}.
\item Added \texttt{\detokenize{tests/test_examples_and_cli.py}} for example subprocess checks, CLI help checks, and a tiny rotated-surface replay output smoke test.
\item Extended CI to exercise installed console scripts, examples, and help output.
\item Validation completed with \texttt{\detokenize{$PYTHON_BIN}} as the active project interpreter: editable install, native build, \texttt{\detokenize{pytest -q}} with \texttt{15 passed}, frontier smoke, \texttt{\detokenize{rotated_surface_d3}} DEM info, \texttt{\detokenize{bravyi_depth7}} DEM info, and all three examples.
\end{itemize}

\subsection*{2026-06-18 Navigation and Public API Orientation}

\begin{itemize}
\item Added \texttt{\detokenize{docs/ARCHITECTURE.md}}, \texttt{\detokenize{docs/COMMANDS.md}}, \texttt{\detokenize{docs/ENVIRONMENT.md}}, \texttt{\detokenize{docs/LICENSING.md}}, \texttt{\detokenize{AGENTS.md}}, a \texttt{\detokenize{Makefile}}, GitHub Actions CI, and tiny runnable examples under \texttt{\detokenize{examples/}}.
\item Added the lightweight \texttt{\detokenize{frontier/}} package with \texttt{\detokenize{frontier.__init__}}, \texttt{\detokenize{frontier.decoder}}, \texttt{\detokenize{frontier.dem}}, and \texttt{\detokenize{py.typed}} as stable public re-exports while keeping existing \texttt{\detokenize{tools.*}} imports and console-script implementations intact.
\item Updated README navigation and \texttt{\detokenize{docs/FILE_SCOPE.md}} so humans and agents can find the architecture guide, command index, environment variables, examples, retained-file audit, and licensing notes quickly.
\item Added module docstrings, argparse help epilogues, and section headers to the retained long \texttt{\detokenize{tools/}} modules without changing decoder math, native extension naming, or public reproduction command semantics.
\item Added lightweight tests for public package imports, console-script module entry points, architecture docs, command docs, and file-scope coverage.
\item Validation completed with a project-local Python interpreter because this shell had no active \texttt{python} alias and the system \texttt{python3} lacked \texttt{setuptools}: native build, \texttt{\detokenize{pytest -q}} with \texttt{11 passed}, frontier smoke with \texttt{\detokenize{--K 16 --Delta 100 --shots 3}}, \texttt{\detokenize{rotated_surface_d3}} DEM info, \texttt{\detokenize{bravyi_depth7}} DEM info, and all three examples.
\end{itemize}

\subsection*{2026-06-15}

\begin{itemize}
\item Created the initial \texttt{frontier} export.
\item Selected repo shape 3 and decoder mode 3: native C++ decoder plus DEM replay/benchmark CLI, with forward \texttt{deadline\_reorder} and backward \texttt{backward\_deadline\_reorder}.
\item Public modes are limited to \texttt{forward\_only}, \texttt{backward\_only}, and \texttt{fwd\_bwd\_committee}.
\item Validation completed: build \texttt{\_frontier\_native}, compile exported Python modules, run \texttt{pytest} with \texttt{3 passed}, and run the smoke benchmark with \texttt{--K 16 --Delta 100 --shots 3}.
\end{itemize}

\subsection*{Open Items}

\begin{itemize}
\item Pushed to GitHub: \texttt{\detokenize{git@github.com:aleverrier/frontier.git}}, branch \texttt{main}; first published commit was \texttt{3b45933}.
\item Continue to keep the standalone export minimal: BB/Gross and surface-code matrix builders, DEM loading, the frontier decoder, replay, benchmark, and tests.
\end{itemize}

\subsection*{2026-06-15 Naming and Matrix Docs Cleanup}

\begin{itemize}
\item Renamed the public exported modules and extension to \texttt{frontier\_*}, including \texttt{\_frontier\_native}, \texttt{frontier\_native.py}, \texttt{tools/frontier\_decoder.py}, \texttt{tools/frontier\_sample\_replay.py}, and \texttt{tools/frontier\_bb144\_benchmark.py}.
\item Renamed public Python API types/functions to \texttt{Frontier*} / \texttt{decode\_frontier*}.
\item Added README matrix-availability notes: static matrix files are not checked in for generated surface-code DEMs; in-repo constructors cover surface-code checks and generated rotated-surface DEMs, while the accepted Gross split-sector DEM builder uses the bundled Gross-code matrix/circuit assets.
\end{itemize}

\subsection*{2026-06-15 Public Path Cleanup}

\begin{itemize}
\item Removed private checkout examples and hardcoded local filesystem defaults from README text, worklogs, and exported Python helpers.
\item Changed optional Gross/Tanner asset handling to require explicit public asset configuration instead of silently falling back to a local developer path.
\item Renamed temporary cache defaults to \texttt{frontier} names.
\end{itemize}

\subsection*{2026-06-15 Repo Scope Cleanup}

\begin{itemize}
\item Audited the standalone export file-by-file for the intended public surface: frontier on BB/Gross and surface-code detector-side matrices.
\item Removed the legacy \texttt{\detokenize{grosscode/decoders/**}} package, including full BP/min-sum, windowed BP/min-sum, local-round, triangle-quotient, triangle small-set-flip, and structure-aware decoder families.
\item Removed old research-only support trees: \texttt{\detokenize{grosscode/bench/**}}, \texttt{\detokenize{grosscode/polar_dem/**}}, projected-location/min-sum helpers, triangle-basis/reference-recovery helpers, Tanner redundant extraction, nonbinary CNOT/quaternary BP helpers, and legacy model/baseline modules.
\item Kept the frontier/native path, BB144/Gross/generalized-bicycle/rotated-surface/surface matrix builders, and split-sector DEM builder.
\item Added \texttt{\detokenize{docs/FILE_SCOPE.md}} as the file-by-file retained-scope audit.
\end{itemize}

\subsection*{2026-06-15 BB144/Gross Reproducibility Docs}

\begin{itemize}
\item Added README instructions for reproducing BB144/Gross split-sector DEM results at a chosen \(p = \texttt{\detokenize{--p-location}}\).
\item Documented matrix inspection via \texttt{\detokenize{frontier-dem-info}} and exact matched full-frame replay via \texttt{\detokenize{frontier-replay}} plus a saved \texttt{\detokenize{sample_rows.csv}}.
\item Clarified that exact published full-frame rows require the same matched sample-row corpus; large sample corpora are not checked into this repo.
\item Made \texttt{\detokenize{frontier-bb144-benchmark --help}} side-effect-free by deferring Matplotlib cache directory creation until after argument parsing.
\end{itemize}

\subsection*{2026-06-15 Minimal Repo Tightening}

\begin{itemize}
\item Replaced the archived BB144/Gross report driver with \texttt{\detokenize{tools/dem_loader.py}}, a small loader and \texttt{\detokenize{frontier-dem-info}} CLI for supported BB/Gross and surface-code detector matrices.
\item Removed archived report, triangle-ordering, and Stim fault-analysis helper files that were not required by the public decoder path.
\item Removed physically dead native C++ code for non-public staged decoding and removed staged replay CSV fields.
\item Made \texttt{\detokenize{frontier-bb144-benchmark}} require an explicit \texttt{\detokenize{--sample-rows}} file instead of defaulting to a local \texttt{\detokenize{results/...}} path.
\item Validation completed after tightening: Python compile with bytecode redirected outside the repo, native build, \texttt{\detokenize{pytest -q -p no:cacheprovider}} with \texttt{3 passed}, frontier smoke with \texttt{\detokenize{--K 16 --Delta 100 --shots 3}}, real \texttt{\detokenize{rotated_surface_d3}} DEM loading through \texttt{\detokenize{tools.dem_loader}}, and help checks for \texttt{\detokenize{tools.dem_loader}} and \texttt{\detokenize{tools.frontier_bb144_benchmark}}.
\end{itemize}

\subsection*{2026-06-15 Fresh-Clone Reproducibility Fixes}

\begin{itemize}
\item Audited a clean clone of \texttt{\detokenize{git@github.com:aleverrier/frontier.git}} as a new user.
\item Fixed the clean Python 3.14 install instructions by installing \texttt{setuptools} and \texttt{wheel} before \texttt{\detokenize{python setup.py build_ext --inplace}}.
\item Replaced public Gross asset configuration with bundled \texttt{\detokenize{grosscode/assets/gross144}} files plus an optional \texttt{\detokenize{GROSSCODE_ASSET_ROOT}} override, and removed the old internal asset-root environment wording from code and docs.
\item Bundled the Gross \(\left[\left[144,12,12\right]\right]\) CSS matrices, BB144/Gross memory X/Z Stim circuits, and materialized \texttt{\detokenize{bravyi_depth7}}, \(p=0.001\) split-sector DEM sparse matrices/priors.
\item Corrected the fresh-clone install command to start with \texttt{\detokenize{python3 -m venv .venv}}, since this macOS host does not provide a \texttt{python} executable before the venv exists.
\item Updated the matrix section to state explicitly that Gross/BB144 static assets are bundled while other supported matrix families are built or generated.
\item Added \texttt{\detokenize{frontier-sample-rows}}, a public CLI that generates \texttt{\detokenize{sample_rows.csv}} from the same detector-side DEM matrices and priors used by \texttt{\detokenize{frontier-replay}}.
\item Documented the complete fresh BB144/Gross DEM reproduction flow at \(p=0.001\), \(\Delta=12\), \(K=512\), and \texttt{\detokenize{fwd_bwd_committee}}: generate 10k matched sample rows, then replay them with the native engine.
\item Made \texttt{\detokenize{frontier-dem-info}} load all requested matrices before printing its CSV header, so missing assets no longer produce a partial CSV.
\item Removed replay summary warnings when pressure diagnostics are disabled.
\item Validation completed: Python compile with bytecode redirected outside the repo, \texttt{\detokenize{pytest -q -p no:cacheprovider}} with \texttt{6 passed}, native build through the fresh-clone venv interpreter, frontier smoke with \texttt{\detokenize{--K 16 --Delta 100 --shots 3}}, \texttt{\detokenize{rotated_surface_d3}} DEM loading, clean one-line missing-assets handling for a bad BB144/Gross \texttt{\detokenize{GROSSCODE_ASSET_ROOT}} override, bundled BB144/Gross matrix loading without an override, and generated-row native replays for both \texttt{\detokenize{rotated_surface_d3}} and a 10-shot BB144/Gross \(p=0.001\), \(\Delta=12\), \(K=512\), \texttt{\detokenize{fwd_bwd_committee}} run.
\end{itemize}

\subsection*{2026-06-15 BB144 Reproduction Success Criteria}

\begin{itemize}
\item Added a README section explaining how a fresh user can tell that the BB144/Gross \(p=0.001\), \(\Delta=12\), \(K=512\), \texttt{\detokenize{fwd_bwd_committee}} reproduction worked.
\item Documented the expected \texttt{\detokenize{frontier-dem-info}} matrix output: \texttt{\detokenize{memory_X}} and \texttt{\detokenize{memory_Z}} detector matrices are \texttt{\detokenize{936x8784}}, logical matrices are \texttt{\detokenize{12x8784}}, with 12 noisy rounds and \texttt{\detokenize{deadline_reorder}}.
\item Documented the expected generated workload shape: 20,000 side rows for 10,000 paired full-frame shots, 40 replay shard tasks when \texttt{\detokenize{--shards-per-side 20}} is used, and a complete native replay recorded in \texttt{\detokenize{run_metadata.json}}.
\item Clarified that \texttt{\detokenize{summary_by_scope.csv}} contains \texttt{\detokenize{memory_X}}, \texttt{\detokenize{memory_Z}}, and \texttt{\detokenize{combined}}, and that the \texttt{\detokenize{combined}} row is the strict full logical FER over paired side rows.
\item Avoided claiming \(\mathrm{FER}=0\): at \(p=0.001\), 10k shots is a smoke-scale reproducibility sample, and no observed failures should be reported as below the resolution of that sample rather than as evidence that the FER is zero.
\item Avoided quoting a wall-clock target in the README. Timing is machine-dependent; the README now asks timing reports to include machine/Python/workers/native availability/batch size and points to transition-evaluation counts for more machine-independent comparison.
\end{itemize}

\subsection*{2026-06-15 Minimal Public Cleanup}

\begin{itemize}
\item Removed the non-self-contained external BB circuit-generation bridge because it depended on a checkout that is not part of this repo.
\item Removed the remaining replay re-decode CLI flags, CSV fields, metadata fields, and report text. The retained replay modes are forward-only, backward-only, and forward/backward committee with the requested \(K\) and \(\Delta\).
\item Replaced the old large progressive helper with \texttt{\detokenize{tools/frontier_progressive.py}}, which contains only the frontier column/layout/order helpers required by the public decoder and DEM loader.
\item Removed an obsolete sample-row option that only applied to the deleted external constructor path.
\item Cleaned docs and worklogs to avoid exact local temporary clone paths while keeping the reproducibility commands and validation notes useful.
\end{itemize}
